SP

### 2018-04-04 SP_data_cleaning.Rmd
All data in, just need polishing!

```{r}
dat <- read.csv("data/spain_clean.csv")
#head(dat)
library(reshape2)
#dat2 <- melt(dat, id.vars = c("locality", "gen_sp", "when"), 
 #            measure.vars = "freq")
#head(dat2, 100)
#sort(unique(dat2$gen_sp))
#clean with clean_sp
dat$gen_sp <- gsub(x = dat$gen_sp, pattern = " ", replacement = "", fixed = TRUE)
#sort(unique(dat$gen_sp))
dat$gen_sp <- gsub(x = dat$gen_sp, pattern = "_", replacement = " ", fixed = TRUE)
#remove sp, NA,
#NA_NA   
#Hylaeus_Variegatus
#source("clean_species.R") #sourced from Evenness project.
library(taxize)
spp <- clean_species(dat$gen_sp, rows = 1)
species <- dat$gen_sp
#table(dat2$gen_sp, dat2$when)
#Andrena_pilipes = carbonaria
#remove A. mellifera
rich_tot <- dcast(dat, when ~ ., value.var = "gen_sp", 
              fun.aggregate = function(x) length(unique(x)))
rich_tot #wow.
#extract genus
pos <- regexpr(dat$gen_sp, pattern = " ", fixed = TRUE)
dat$genus <- substr(dat$gen_sp, 
                    start = 1,
                    stop = pos-1)
genus_tot <- dcast(dat, when ~ ., value.var = "genus", 
              fun.aggregate = function(x) length(unique(x)))
genus_tot2 <- dcast(dat,  genus ~ when, value.var = "genus", 
              fun.aggregate = function(x) length(unique(x)))
genus_tot #wow, reversed??
genus_tot2
subset(genus_tot2, now == 0 & asensio == 1)
subset(genus_tot2, now == 1 & asensio == 1)
subset(genus_tot2, now == 1 & asensio == 0)

#remove non apis genus! check with Curro.

@@

#calculate richness per site and time period
#head(dat)
rich <- dcast(dat, locality ~ when, value.var = "gen_sp", 
              fun.aggregate = function(x) length(unique(x)))
#rich
#t.test(rich$asensio, rich$now, paired = TRUE)

@@

library(vegan)
temp <- subset(dat, when == "now")
comm <- dcast(temp, locality ~ gen_sp, value.var = "freq", 
              fun.aggregate = sum, na.rm = TRUE)
rownames(comm)[-8] <- as.character(comm$locality[-8])
comm2 <- comm[,-1] 
#rowSums(comm2) #toro is low, but same sampling effort. 
temp <- specaccum(comm2, "random")
#plot(temp)
#rarecurve(comm2)
now <- rarefy(x = comm2, sample = 120)

temp <- subset(dat, when == "asensio")
comm <- dcast(temp, locality ~ gen_sp, value.var = "freq", 
              fun.aggregate = sum, na.rm = TRUE)
#comm <- dcast(temp, locality ~ gen_sp, value.var = "freq", 
 #             fun.aggregate = length, na.rm = TRUE)
rownames(comm)[-8] <- as.character(comm$locality[-8])
comm2 <- comm[,-1]
#rowSums(comm2) #sanbernardo useless. Others ok! Delives super sampled!!
temp <- specaccum(comm2, "random")
#plot(temp)
#rarecurve(comm2) #for delives quite complete!
asensio <- rarefy(x = comm2, sample = 120)

ttest <- t.test(asensio[c(1,2,3,5,6,7)], now[c(1,2,3,5,6,7)], paired = TRUE)

boxplot(asensio[c(1,2,3,5,6,7)], now[c(1,2,3,5,6,7)], names = c("1980's", "2016"), ylab = "rarefied species richness", 
            las = 1, border = "light grey")
points(x = c(rep(1,6), rep (2,6)), y = c(asensio[c(1,2,3,5,6,7)], now[c(1,2,3,5,6,7)]), col = rep(c(1,2,3,5,6,7),2), pch = 19)
for(i in c(1,2,3,5,6,7)){
    lines(x = 1:2, y = c(asensio[i], now[i]), col = i)
}

#do by coverage?
#clean sp.
```

>>>>

```{r}
dat <- read.csv("data/spain_clean.csv")
#head(dat)
library(reshape2)
#dat2 <- melt(dat, id.vars = c("locality", "gen_sp", "when"), 
 #            measure.vars = "freq")
#head(dat2, 100)
#sort(unique(dat2$gen_sp))
#clean with clean_sp
dat$gen_sp <- gsub(x = dat$gen_sp, pattern = " ", replacement = "", fixed = TRUE)
#sort(unique(dat$gen_sp))
dat$gen_sp <- gsub(x = dat$gen_sp, pattern = "_", replacement = " ", fixed = TRUE)
#remove sp, NA,
#NA_NA   
#Hylaeus_Variegatus
#source("clean_species.R") #sourced from Evenness project.
library(taxize)
#SLOW!!
spp <- clean_species(dat$gen_sp, rows = 1) 
spp[which(as.character(spp$species) != as.character(spp$final_names)),]
spp$species
temp <- spp[which(is.na(spp$final_names)),]
str(temp)
options(max.print = 4000)
print(temp)
spp[which(spp$species == "Andrena pilipes"),2] <- "Andrena carbonaria"
spp[which(spp$species == "Megabombus laesus"),2] <- "Bombus laesus"
spp[which(spp$species == "Colletes eousMorice,1904"),2] <- "Colletes eous"
spp[which(spp$species == "Reanthidium nigricolle"),2] <- "Anthidium nigricolle"
spp[which(spp$species == "Megabombus muscorum"),2] <- "Bombus muscorum"
spp[which(spp$species == "Megabombus ruderatus"),2] <- "Bombus ruderatus"
spp[which(spp$species == "Megabombus sylvarum"),2] <- "Bombus sylvarum"

head(dat)
head(spp)
dat <- merge(dat, spp[,1:2], by.x = "gen_sp", by.y = "species", all.x = TRUE)
head(dat, 100)
#remove A. mellifera
dat <- subset(dat, !matched_name2 %in% c("Apis mellifera"))
# and genus only
pos <- regexpr(dat$matched_name2, pattern = " ", fixed = TRUE)
str(pos)
dat2 <- dat[-which(pos < 1),]
dat2$matched_name2
#NA
dat3<- dat2[-which(is.na(dat2$matched_name2)),]
unique(dat3$matched_name2)
dat <- dat3

rich_tot <- dcast(dat, when ~ ., value.var = "matched_name2", 
              fun.aggregate = function(x) length(unique(x)))
rich_tot #wow.
#extract genus
pos <- regexpr(dat$matched_name2, pattern = " ", fixed = TRUE)
dat$genus <- substr(dat$matched_name2, 
                    start = 1,
                    stop = pos-1)
genus_tot <- dcast(dat, when ~ ., value.var = "genus", 
              fun.aggregate = function(x) length(unique(x)))
genus_tot2 <- dcast(dat,  genus ~ when, value.var = "genus", 
              fun.aggregate = function(x) length(unique(x)))
genus_tot #wow, reversed??
genus_tot2
subset(genus_tot2, now == 0 & asensio == 1)
subset(genus_tot2, now == 1 & asensio == 1)
(x <- subset(genus_tot2, now == 1 & asensio == 0))

#remove non apis genus! check with Curro.
#Anthocopa is Hoplitis
#Chalicodoma
#Pseudapis

dat4 <- dat[which(!dat$genus %in% x$genus[c(1:5,7:25,27:34)]),]
dat <- dat4
#Heliophila <-  Anthophora
#Metallinella <- Osmia
#Reanthidium <-  Anthidium
head(dat)
dat[which(dat$genus == "Heliophila"),"matched_name2"] <- "Anthophora bimaculata"
dat[which(dat$genus == "Metallinella"),"matched_name2"] <- "Osmia brevicornis"
dat[which(dat$genus == "Reanthidium"),"matched_name2"] <- "Pseudoanthidium reticulatum"

#resave data
write.csv(dat, "data/spain_clean_used.csv", row.names = F)

dat <- read.csv("data/spain_clean.csv")
#Repeat richness and genus
rich_tot <- dcast(dat, when ~ ., value.var = "matched_name2", 
              fun.aggregate = function(x) length(unique(x)))
rich_tot #wow.
#extract genus
pos <- regexpr(dat$matched_name2, pattern = " ", fixed = TRUE)
dat$genus <- substr(dat$matched_name2, 
                    start = 1,
                    stop = pos-1)
genus_tot <- dcast(dat, when ~ ., value.var = "genus", 
              fun.aggregate = function(x) length(unique(x)))
genus_tot2 <- dcast(dat,  genus ~ when, value.var = "genus", 
              fun.aggregate = function(x) length(unique(x)))
genus_tot #wow, lot's of genus lost!
genus_tot2
subset(genus_tot2, now == 0 & asensio == 1) #20 lost!
subset(genus_tot2, now == 1 & asensio == 1) #27 stable
subset(genus_tot2, now == 1 & asensio == 0) #2 added!
genus_tot3 <- dcast(dat,  genus ~ when, value.var = "matched_name2", 
              fun.aggregate = function(x) length(unique(x)))
genus_tot3
colSums(genus_tot3[,2:3])

@@

#calculate richness per site and time period
#head(dat)
rich <- dcast(dat, locality ~ when, value.var = "matched_name2", 
              fun.aggregate = function(x) length(unique(x)))
(rich <- rich[-8,])
t.test(rich$asensio[-4], rich$now[-4], paired = TRUE) #0.056!

@@

library(vegan)
str(dat)
unique(dat$freq)
dat[which(is.na(dat$freq)),"freq"] <- 1 
dat$freq <- as.numeric(dat$freq)
temp <- subset(dat, when == "now")
comm <- dcast(temp, locality ~ matched_name2, value.var = "freq", 
              fun.aggregate = sum, na.rm = TRUE)
rownames(comm)[-8] <- as.character(comm$locality[-8])
comm2 <- comm[,-1] 
comm2 <- comm2[-8,] 
rowSums(comm2) #toro is low, but same sampling effort. 
plot(specaccum(comm2, "random"))
rarecurve(comm2) #Good!
now <- rarefy(x = comm2, sample = 1000)

temp <- subset(dat, when == "asensio")
comm <- dcast(temp, locality ~ matched_name2, value.var = "freq", 
              fun.aggregate = sum, na.rm = TRUE)
rownames(comm)[-8] <- as.character(comm$locality[-8])
comm2 <- comm[,-1]
rowSums(comm2) #sanbernardo useless. Others ok! Delives super sampled!!
plot(specaccum(comm2, "random"))
rarecurve(comm2) #for delives quite complete!
head(comm2)
rarecurve(comm2[-1,]) #for the rest is ok, but sanbernardo.
abline(v = 1000)
asensio <- rarefy(x = comm2, sample = 1000)

ttest <- t.test(asensio[c(1,2,3,5,6,7)], now[c(1,2,3,5,6,7)], paired = TRUE)
ttest #0.05828

boxplot(asensio[c(1,2,3,5,6,7)], now[c(1,2,3,5,6,7)], names = c("1980's", "2016"), ylab = "rarefied species richness", 
            las = 1, border = "light grey")
points(x = c(rep(1,6), rep (2,6)), y = c(asensio[c(1,2,3,5,6,7)], now[c(1,2,3,5,6,7)]), col = rep(c(1,2,3,5,6,7),2), pch = 19)
for(i in c(1,2,3,5,6,7)){
    lines(x = 1:2, y = c(asensio[i], now[i]), col = i)
}
```
